Predict Models

### 2024-03-15 scripts/classify_patches.Rmd
creating surfaces for predicting phenology onto
arunning a second round, accidentally wrote over some original sections

lapply(f, patcheR) # roughly 72 hours.

@@

@@

lapply(f, FUN = landscapR, pout = 'patch_metrics', pRout = 'fine_patches')

>>>>

lapply(f, patcheR) # roughly 72 hours. 
```

```{r coarse patches 2, eval = F}

targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>%

  ## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

target_taxa <- c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')

## these species missed, fixed on March 13th
target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, target_taxa)) %>% 
  pull(taxon)

f <- file.path(p, list.files(p, pattern = '1k-'))
files <- f[grep(paste(paste0(gsub(' ', '_', target_taxa)), collapse = '|'), basename(f))]

rm(target_taxa)

lapply(files, patcheR) # roughly 5 hours. 
```

@@

targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>%

  ## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

target_taxa <- c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')

## these species missed, fixed on March 13th
target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, target_taxa)) %>% 
  pull(taxon)

files <- f[grep(paste(paste0(gsub(' ', '_', target_taxa)), collapse = '|'), basename(f))]

@@

lapply(files, FUN = landscapR, pout = 'patch_metrics', pRout = 'fine_patches')

### 2024-03-15 scripts/predict_models.Rmd
aggregate cells before finding patches
arunning a second round, accidentally wrote over some original sections

pmodels <- '../results/rf_models'
f <- file.path(pmodels, list.files(pmodels))
f <- data.frame(
  path = f,  
  species = gsub('-.*$', '', basename(f)), 
  time = gsub('[A-Za-z]', '', basename(f))
  ) |>
  mutate(time = gsub('_-|[.]', '', time),
         time = gsub('_', ' ', time), 
         time = trimws(time),
         time = gsub('^.1-|^1-', '', time), 
         time = as.POSIXct(time)) |>
 # group_by(species) |>
  filter(str_detect(species, '1k$')) |>
 # slice_max(order_by = time, n = 1) #|>
  pull(path)

occpath <- '../data/raw/occurrence/combined_records'
occ_files <- list.files(occpath, pattern = 'shp$')

models_occ <- data.frame(
  path2model = f, occurrence_data = occ_files
) %>% 
  mutate(
    species = gsub('-.*$', '', basename(path2model)),
    match = if_else(species == gsub('[.]shp$', '', occurrence_data), T, F )
  ) # all look good!

@@

lapply(models_occ[68:335], predict_wrapper)

```

>>>>

targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>%

  ## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

target_taxa <- c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')

## these species missed, fixed on March 13th
target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, target_taxa)) %>% 
  pull(taxon)

pmodels <- '../results/rf_models'
f <- file.path(pmodels, list.files(pmodels))
files <- f[grep(paste(paste0(gsub(' ', '_', target_taxa)), collapse = '|'), basename(f))]

rm(target_taxa)

occpath <- '../data/raw/occurrence/combined_records'

occ_files <- paste0( basename(gsub('1k.*', '', files)), '.shp')

models_occ <- data.frame(
  path2model = files, occurrence_data = occ_files
) %>% 
  mutate(
    species = gsub('-.*$', '', basename(path2model)),
    match = if_else(species == gsub('[.]shp$', '', occurrence_data), T, F )
  ) # all look good!

@@

lapply(models_occ, predict_wrapper)
```

```{r predict second set of models}

targets <- read.csv('../data/raw/2024sos_species.csv', na.strings = "") %>%

  ## we will not model a few species which are widely abundant, hence do not require modelling, and would utilize enormous computational resources.
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon))

target_taxa <- c('Brodiaea elegans', 'Krameria bicolor', 'Lotus humistratus', 'Adenostoma fasciculatum')

## these species missed, fixed on March 13th
target_taxa <- read.csv('../data/raw/2024sos_species-FIXED.csv', na.strings = "") %>% 
  filter(!species %in% c('tridentata', 'millefolium', 'dioica', 'negundo')) %>%  # purshia,
  unite('taxon', genus:infraspecies, na.rm = T, sep = ' ') %>% # artemisia, and larrea !
  mutate(taxon = str_squish(taxon)) %>% 
  filter(! taxon %in% c(targets$taxon, target_taxa)) %>% 
  pull(taxon)

pmodels <- '../results/rf_models'
f <- file.path(pmodels, list.files(pmodels))
files <- f[grep(paste(paste0(gsub(' ', '_', target_taxa)), collapse = '|'), basename(f))]

rm(target_taxa)

occpath <- '../data/raw/occurrence/combined_records'

occ_files <- paste0( basename(gsub('1k.*', '', files)), '.shp')

models_occ <- data.frame(
  path2model = files, occurrence_data = occ_files
) %>% 
  mutate(
    species = gsub('-.*$', '', basename(path2model)),
    match = if_else(species == gsub('[.]shp$', '', occurrence_data), T, F )
  ) # all look good!

lapply(models_occ, predict_wrapper)
```
